Plots with ggplot2
------------------
We can use the ggplot2 package to make visual summaries of the 
values stored in a data frame.  To do this, we must fist load the 
ggplot2 package by running the command
```{r}
library("ggplot2")
```

To make a histogram, we use the command
```{r}
ggplot(bikedata, aes(passing.distance)) + geom_histogram()
```
This command involves a many subcomponents, whose meanings are
as follow:

  * The `ggplot` function creates a new *plot object*.
  
  * The first argument `bikedata` is the data frame that stores
    the values to use in the plot.
    
  * The second argument, `aes(passing.distance)` specifies the
    *aesthetic mappings* between the data frame variables and the
    plot object.  In this case, `passing.distance` gets mapped
    to the plot variable.
    
  * By itself, the subcommand
    `ggplot2(bikedata, aes(passing.distance))`
    creates an empty plot object where `passing.distance` is the
    plot variable.  This subcommand does not produce any output.
    
  * The subcommand `+ geom_histogram()` means
    "add a histogram geometric object (geom) to the plot object".

We specify with bin width, and color, change the axis labels, and 
add a title using the command
```{r, tidy=FALSE}
(ggplot(bikedata, aes(passing.distance))
  + geom_histogram(binwidth=0.1, color="darkgreen", fill="white")
  + xlab("Passing Distance")
  + ylab("Count")
  + ggtitle("Histogram of Passing Distance"))
```


See the [ggplot documentation](http://docs.ggplot2.org/current) for many other ways to customize your plots.

