Introduction to R
=================

Preparing Data
--------------
Believe it or not, reading in data is often the hardest part of 
working with R.  If you collect and store your data in Excel or 
Google Docs, you will need to carefully format your spreadsheet.
It should obey the following rules:

 1. The spreadsheet should contain a single sheet.
 
 2. Row 1 should contain variable names in consecutive cells, 
    starting with Cell A1.  For convenience, the names should
    be comprised of lowercase words and contain no symbols or
    punctuation.
 
 3. Subsequent rows (starting with Row 2) should contain your 
    observations (data). 

 4. If the value of a variable is missing for a particular 
    observation, the coressponding cell in the spreadsheet should 
    be empty.

 5. All other cells in the spreadsheet should be empty.
 
This sounds straightforward, but many spreadsheets that you find 
"in the wild" do not obey these rules.  You will have to reformat 
these spreadsheets, usually by deleting empty rows and columns
and by deleting notes and other annotations.


Even if your spreadsheet is formatted as above, R cannot open
Excel files.  To save your data to an R-compatible file format, 
exportyour data as a "Column-Separated Value" (CSV) file.  You can 
do this from the File menu in Excel.


Reading Data into R
-------------------
If you have a property-formatted CSV file, you can read it into R 
using the `read.csv` function.  There are two ways to specify the 
file.  To use your system's file chooser, run the command.
```{r, eval=FALSE}
    data <- read.csv(file.choose())
```    
Alternatively, if you know the name of the file, you can pass it
directly to the `read.csv` function.  Note that if you pass the 
file name directly, you must either specify the full path to the 
file, or you must set the "working directory" to be the directory 
thatcontains the file.  To set the working directory, either use 
the `setwd` function or run the *Set Working Directory* command 
from RStudio's **Session** menu.

Suppose that I want to open a file named "bikedata.csv", which is 
stored in the "~/Datasets" directory on my system.  Ifirst set the 
working directory to "~/Datasets" by Choosing **Session** > *Set 
Working Directory* > *Choose Directory ...*.  This will execute
the command
```{r, eval=FALSE}
setwd("~/Datasets")
```
(In fact, if I do not want to use the menu system, then I can just
type this command directly to achieve the same effect.)  Once the
working directory is set, I can read in the data to a variable
named `data` by executing the command
```{r}
bikedata <- read.csv("bikedata.csv")
```


Variables
---------
In R, we use the term "variable" to refer to a name-value pair. 
You should not confuse this concept with the types of variables
you have seen in your math classes (they are similar in some ways, 
but different in others).


In the last section, when we ran the command
`data <- read.csv(file.choose())` we created a variable with name 
`data` and value equal to the contents of the chosen file.


To create a variable or to assign a new value to an existing 
variable, use the assignment command (`<-`),  which is meant to
look like an arrow pointing from the value to the variable name.
For example, the command
```{r}
a <- 2.7
```
means "assign the value `2.7` to the variable named `a`".  
Another way to read this is "variable `a` gets the value `2.7`".


When you have a variable, you can use the name in place of
the value:
```{r}
a + 10
5 * a
```
You can see the value of a variable, by typing its name and 
pressing enter:
```{r}
a
```


An alternative way to read in a file to the variable named `data` 
is to run the following sequence of commands:
```{r, eval=FALSE}
filename <- file.choose()
data <- read.csv(filename)
```
The first command asks the user to choose a file, and stores the 
resulting name in the `filename` variable.  This variable contains 
the name of the file, but not the actual contents.  The second 
command takes the name of the file, opens it, reads the contents 
into memory, and stores the result in the `data` variable.


Functions
---------
Besides variables, the other main concept you need to learn in
R is that of a *function*.  You are probably familiar with the 
concept of a function from your mathematics courses, and a 
function in R is very similar: a function is something that takes zero or more values, then performs a sequence of actions
and returns a result.


We have already seen three functions: `file.choose`, `read.csv`,
and `setwd`.  We *call* a function by putting a pair of parenteses
`()` after the function name.  Many functions, including
`read.csv` and `setwd` require one or more values as input.  We 
refer to these values as *arguments*, and we specify them by
putting the values inside the parentheses.  When we do so, we
say that we are *passing the value* of the argument to the 
function.


Sometimes a function will have optional arguments.  These are
arguments that, if left unspecified, will be given reasonable
default values.  For example, by default, the `file.choose`
function forces the user to choose an existing file.  To allow
the user to choose a name for new file, pass the argument
`new=TRUE` to file `file.choose` function:
```{r, eval=FALSE}
file.choose(new=TRUE)
```
Before, we did not specify the `new` argument, and it defaulted
to the value `FALSE`.


Data Frames
-----------
The `read.csv` command opens the file and reads the data into a 
type of object called a ``dataframe''.  Conceptually, a data frame 
is just like a spreadsheet: it has columns, corresponding to 
variables, and rows, corresponding to observations.  Each row and 
column has a name.  Usually, the row names are the character
strings "1", "2", etc., but this is not always the case.


To see the first 6 rows in the `bikedata` data frame, run the 
command
```{r}
head(bikedata)
```
We can see that there are nine columns, named `vehicle`,
`colour`, `passing.distance`, `street`, `helmet`, `kerb`,
`datetime`, `bikelane`, and `city`.


To see a summary of the entire data frame, use the ``summary`` function:
```{r}
summary(bikedata)
```


Extracting Columns
------------------
Let's say we want to investigate the `passing.distance` variable.  
To do this, we must first *extract* that column from the
`bikedata` dataframe.  There are three ways to do this:
```{r}
x <- bikedata$passing.distance
x <- bikedata[["passing.distance"]]
x <- bikedata[,"passing.distance"]
```
All three commands are equivalent ways to extract the
`passing.distance` column and store it in a variable named `x`.   
The `$` form is the most common, but you will sometimes see the 
other two forms, as well.


Vectors
-------
Data frame columns are stored in a data type called a "vector".
Conceptually, a vector is a one-dimensional array of values,
indexed by integers starting at `1`.  Most functions in R operate 
on vectors.


You can access individual values by using  double square-brackets.  
For example, to see the first element of the vector, type the
command
```{r}
x[[1]]
```
To see the fifth value, type the command
```{r}
x[[5]]
```
To see how many elements are contained in the vector, use the
`length` function:
```{r}
length(x)
```
To see the last element, type
```{r}
x[[length(x)]]
```


To extract a subvector, use single square brackets.  For example,
the subvector consisting of the first 25 elements is
```{r}
x[1:25]
```
Here, `1:25` is shorthand for "integers 1 to 25".  Since not all 
25 values fit onto a single line, R wraps the values.  At the 
start of each line, R prints the index of the first value on the 
line in square brackets.  Looking at the output above, we can
see that `1.410` is the 12th element and `1.492` is the 24th 
element of the result.


You may have asked yourself earlier why the output of `x[[1]]`
and other similar commands was prefixed by `[1]`.  The reason for 
this is that R doesn't have the concept of a "single value" or 
"scalar".  The only way to represent the value of `x[[1]]` is
as a length-one vector.  The output

    [1] 2.114

denotes a vector with a single value (`2.114`), stored at
index `1`.


Since there is no concept of a "scalar" in R, the command
`x[[1]]` is equivalent to `x[1:1]`, which is also the same as
`x[1]`.  In other programming languages, `x[[1]]`,
"the first element of `x`", and `x[1]`, "the subvector of `x` 
starting and ending at index `1`" would be different; in R, these 
are identical.  Because of this, most people use single brackets instead of double brackets when indexing vectors, writing
`x[1]` and `x[5]` instead of `x[[1]]` and `x[[5]]]`.


Descriptive Statistics for Numeric Variables
--------------------------------------------
There are a variety of functions for computing descriptive
statistics for the values stored in a vector.

  * Sum of values:
    ```{r}
    sum(x)
    ```

  * Measures of central tendency (sample mean and median):
    ```{r}
    mean(x)
    median(x)
    ```

  * Measures of variability (sample standard deviation and sample 
    variance):
    ```{r}
    sd(x)
    var(x)
    ```

  * Extreme values (minimum and maximum):
    ```{r}
    min(x)
    max(x)
    ```

  * Quantiles:
    ```{r}
    quantile(x, .25) # first quartile
    quantile(x, .75) # third quartile
    quantile(x, .99) # 99th percentile
    ```


Plots for Numeric Variables
---------------------------
We can use the `hist` command to make a histogram of the values 
stored in a vector:
```{r}
hist(x)
```
By default, the output looks fine when printed in black and white, 
but it isn't very pretty.  We can specify the bin color, change the axis lables, and omit the main title by passing additional arguments to this `hist` function
```{r, tidy=FALSE}
hist(x, col="steelblue", xlab="Passing Distance", ylab="Count", 
     main="")
```


Use the `boxplot` and `qqnorm` commands to make boxplots and
normal probability plots, as in the following examples:
```{r, tidy=FALSE}
boxplot(x, border="darkred", ylab="Passing Distance")
qqnorm(x, col="darkgreen", xlab="Normal Quantiles",       
       ylab="Passing Distance Quantiles",
       main="")
```


Categorical Variables
---------------------
So far, we have seen how to use R to summarize and plot a numeric 
(quantitative) variable.  R also has very good support for
categorical (qualitative) variables, referred to as *factors*.


To see levels, the set of possible values for a factor variable,
use the `levels` function.  For example, to see the levels of
the `colour` variable:
```{r}
levels(bikedata$colour)
```


To tabulate the values of the variable, use the `table` command,
as in
```{r}
table(bikedata$colour)
````
Note: be default, the `table` command omits missing values.  To 
include these values in the output, include `useNA="ifany"` in
the call to `table`:
```{r}
table(bikedata$colour, useNA="ifany")
```


We can present tabulated counts in a bar plot using the following
commands
```{r}
tab <- table(bikedata$colour, useNA="ifany")
barplot(tab)
```
Usually, it makes sense to arrange the table values in decreasing 
order.  Here is an example with sorted counts that adds axis 
labels and changes the bar colors:
```{r, tidy=FALSE}
barplot(sort(tab, decreasing=TRUE), xlab="Colour", ylab="Count",
        col="steelblue")
```


Inference for a Population Mean
-------------------------------
We can use the `t.test` function to test a hypothesis
about a population mean.
```{r}
t.test(bikedata$passing.distance)
```
This reports the t statistic, the degrees of freedom, the p-value, 
and the sample mean.  The command also reports a 95% confidence 
interval for the population mean.  To change the confidence level, 
use the `conf.level` argument, as in
```{r}
t.test(bikedata$passing.distance, conf.level=0.99)
```


By default, the null hypothesis is that the true (population)
mean is equal to 0, and the alternative hypothesis is that the
true mean is not equal to 0.  To use a different null, pass the
`mu` argument.  To use a different alternative, pass
`alternative="less"` or `alternative="greater"`.  For example, to
test the null hypothesis that the true mean is equal to 1.5
against the alternative that it is greater, run the command
```{r}
t.test(bikedata$passing.distance, alternative="greater", mu=1.5)
```
Note that for a one-sided alternative, the confidence interval
is one-sided, as well.


Inference for a Population Proportion
-------------------------------------
To perform a test on a population proportion, use the `prop.test` 
function.  This performs a test on a population proportion that is 
slightly different than the one we cover in the core statistics 
course, but it will give you a very similar answer.


In the first argument, specify `x`, the number of successes; in 
the second argument, specify `n`, the number of trials.  By 
default, the null value of the population proportion is `0.5`; to 
specify a different value, use the `p` argument.


For example, to test the null hypothesis that the true proportion 
of cars passing the rider on his route is exactly equal to 40%,
we first tabulate the `colour` variable,
```{r}
table(bikedata$colour)
```
In this instance, the number of "successes" is equal to the number 
of blue cars, `636`.  Recall that some of the values for the
`colour` variable are missing.  If the missingness is unrelated to 
the actual color, then we can safely ifnore these values; in this 
case, the number of "trials" is equal to the sumof the counts for 
all of the non-missing values.
```{r}
sum(table(bikedata$colour))
```
Now, to test the proportion, we run the command:
```{r}
prop.test(636, 2341, p=0.40)
```


As with the `t.test` function, we can use a one-sided alternative
or specify a different confidence level for the interval by
using the `alternative` or `conf.level` argument, respectively.
Here is a test of the null that the true proportion is equal to
`0.5` against the alternative that it is less, along with a
one-sided 99% confidence interval:
```{r}
prop.test(636, 2341, p=0.50, alternative="less", conf.level=0.99)
```
