---
title: Forecasting Time Series
subtitle: Unit Root Tests & Cointegration
author: Patrick Perry 
date: May 2, 2017
---

```{r setup, cache=FALSE, echo=FALSE}
library("RColorBrewer")    # brewer.pal
library("knitr")           # opts_chunk

# terminal output
options(width = 80)

# color palette
palette(brewer.pal(6, "Set1"))

# RNG seed
set.seed(0)

# code chunk options
opts_chunk$set(cache=TRUE, fig.align="center", comment=NA, echo=TRUE,
               tidy=FALSE)
```

Preamble
--------

This lecture is based on Cowpertwait & Metcalfe,
*Introductory Time Series with R*, Chapter 11.

```{r}
library("forecast")
library("tseries")
```


Exchange Rate Data
------------------

```{r}
# Daily exchange rates for UK pounds and the Euro, 
# from January 2004 to December 2007, per US dollar.
#
xrates <- read.table("http://www.maths.adelaide.edu.au/andrew.metcalfe/Data/us_rates.dat",
                     header=TRUE)
n <- nrow(xrates)
time <- 1:n

euro <- xrates$EU
gbp <- xrates$UK
```


Time Series Plots
-----------------

```{r}
ylim <- range(euro, gbp)
plot(time, euro, ylim=ylim, type="l", col=2, ylab="Exchange Rate", xlab="Time")
lines(time, gbp, col=3)
legend("topright", inset=0.025, legend=c("euro", "gbp"), col=c(2, 3), lty=2)
```


ACFs, PACFs for Euro
--------------------

**Euro**

```{r}
par(mfrow=c(1,2))
Acf(euro)
Pacf(euro)
```

**Diff. Euro**

```{r}
par(mfrow=c(1,2))
Acf(diff(euro))
Pacf(diff(euro))
```

**2nd Diff. Euro**

```{r}
par(mfrow=c(1,2))
Acf(diff(diff(euro)))
Pacf(diff(diff(euro)))
```


ACFs, PACFs for GBP
-------------------


**Great Britain Pound**

```{r}
par(mfrow=c(1,2))
Acf(gbp)
Pacf(gbp)
```


**Diff. Great Britain Pound**

```{r}
par(mfrow=c(1,2))
Acf(diff(gbp))
Pacf(diff(gbp))
```

**2nd Diff. Great Britain Pound**

```{r}
par(mfrow=c(1,2))
Acf(diff(diff(gbp)))
Pacf(diff(diff(gbp)))
```


Dickey-Fuller Test for Euro
---------------------------

**Step 1: Least Squares Fit**

```{r}
lag.euro <- c(NA, euro)[1:length(euro)]
lsfit.euro <- lm(euro ~ lag.euro)
summary(lsfit.euro)
```


**Step 2: Compute Test Statistic**

```{r}
(tau.mu.euro <- (0.997650 - 1) / (0.003219))
```


**Step 3: Perform Test**

Use value of n to find appropriate row of Table 8.5.2:

```{r}
print(n)
```

With n > 1000, use the "n = 500" or "n = Infinity" row of the table.
Since -2.57 < tau.mu < -0.44, it follows that 0.10 < p < 0.90.  We do not
reject the null hypothesis; we do not have evidence that euro is *not* a
random walk.


Dickey-Fuller Test for GBP
--------------------------


**Step 1: Least Squares Fit**

```{r}
lag.gbp <- c(NA, gbp)[1:length(gbp)]
lsfit.gbp <- lm(gbp ~ lag.gbp)
summary(lsfit.gbp)
```


**Step 2: Compute Test Statistic**

```{r}
(tau.mu.gbp <- (0.994113 - 1) / (0.003510))
```


**Step 3: Perform Test**

Use value of n to find appropriate row of Table 8.5.2:

```{r}
print(n)
```

As with Euro, since -2.57 < tau.mu < -0.44, it follows that 0.10 < p < 0.90.
We do not reject the null hypothesis; we do not have evidence that GBP is
*not* a random walk.


Ad-Hoc Test for Cointegration
-----------------------------

**Least Squares Fit**

```{r}
fit <- lm(euro ~ gbp)
summary(fit)
```

**Fit Residuals vs. Lag**

```{r}
resid <- residuals(fit)
lag.resid <- c(NA, resid)[1:length(resid)]
fit.resid <- lm(resid ~ lag.resid)
summary(fit.resid)
```


**Perform DF Test on Residuals**
```{r}
(tau.mu.resid <- ((9.846e-01) - 1) / (6.503e-03))
```

Since -2.57 < tau.mu < -0.44, we have 0.10 < p < 0.90.  We cannot reject the
null that the two series are not cointegrated.  We do not have evidence of
co-integration.


Phillips-Ouliaris Test for Cointegration
----------------------------------------

```{r}
po.test(cbind(euro, gbp))
```

Note: test is sensitive to order (also true for ad-hoc procedure above).

```{r}
po.test(cbind(gbp, euro))
```

Conclusion: some weak evidence of co-integration.




Residual Plots
--------------

**Time Series Plot**

```{r}
plot(time, resid, type="l")
```

**ACF, PACF of Residuals**

```{r}
par(mfrow=c(1,2))
Acf(resid)
Pacf(resid)
```

