---
title: Forecasting Time Series
subtitle: Analysis of the Google Series
author: Patrick Perry 
date: March 7, 2017
---

```{r setup, cache=FALSE, echo=FALSE}
library("RColorBrewer")    # brewer.pal
library("knitr")           # opts_chunk

# terminal output
options(width = 80)

# color palette
palette(brewer.pal(6, "Set1"))

# set random number generator seed to ensure consistent runs
set.seed(0)

# code chunk options
opts_chunk$set(cache=TRUE, fig.align="center", comment=NA, echo=TRUE,
               tidy=FALSE)
```



---

```{r}
library("forecast")
```


Google Series (GOOG)
--------------------
Daily closing price: May 2, 2005 to December 29, 2006.

```{r}
data <- read.csv("http://ptrckprry.com/course/forecasting/data/google.csv")
GOOG <- data$google
n <- length(GOOG)
time <- 1:n
plot(time, GOOG, type="l", col=2)
```


Log GOOG
--------

```{r}
log.GOOG <- log(GOOG)
plot(time, log.GOOG, type="l", col=2)
Acf(log.GOOG)
Pacf(log.GOOG)
```


Diff Log GOOG
-------------

```{r}
diff.log.GOOG <- c(NA, diff(log.GOOG))

plot(time, diff.log.GOOG, type="l", col=2)
Acf(diff.log.GOOG)
Pacf(diff.log.GOOG)
```


2nd Diff Log GOOG
-----------------

```{r}
diff2.log.GOOG <- c(NA, diff(diff.log.GOOG))

plot(time, diff2.log.GOOG, type="l", col=2)
Acf(diff2.log.GOOG)
Pacf(diff2.log.GOOG)
```


ARIMA(0, 1, 0)
--------------

```{r}
fit.00 <- Arima(log.GOOG, c(0, 1, 0), include.constant=FALSE)
print(fit.00)
fit.00$aicc
```


AICC for Candidate Models (Buggy Version)
-----------------------------------------

```{r}
# Without constant:
fit.00 <- Arima(log.GOOG, c(0, 1, 0), include.constant=FALSE)
fit.01 <- Arima(log.GOOG, c(0, 1, 1), include.constant=FALSE)
fit.02 <- Arima(log.GOOG, c(0, 1, 2), include.constant=FALSE)
fit.10 <- Arima(log.GOOG, c(1, 1, 0), include.constant=FALSE)
fit.11 <- Arima(log.GOOG, c(1, 1, 1), include.constant=FALSE)
fit.12 <- Arima(log.GOOG, c(1, 1, 2), include.constant=FALSE)
fit.20 <- Arima(log.GOOG, c(2, 1, 0), include.constant=FALSE)
fit.21 <- Arima(log.GOOG, c(2, 1, 1), include.constant=FALSE)
fit.22 <- Arima(log.GOOG, c(2, 1, 2), include.constant=FALSE)

# With constant:
fit.00c <- Arima(log.GOOG, c(0, 1, 0), include.constant=TRUE)
fit.01c <- Arima(log.GOOG, c(0, 1, 1), include.constant=TRUE)
fit.02c <- Arima(log.GOOG, c(0, 1, 2), include.constant=TRUE)
fit.10c <- Arima(log.GOOG, c(1, 1, 0), include.constant=TRUE)
fit.11c <- Arima(log.GOOG, c(1, 1, 1), include.constant=TRUE)
fit.12c <- Arima(log.GOOG, c(1, 1, 2), include.constant=TRUE)
fit.20c <- Arima(log.GOOG, c(2, 1, 0), include.constant=TRUE)
fit.21c <- Arima(log.GOOG, c(2, 1, 1), include.constant=TRUE)
fit.22c <- Arima(log.GOOG, c(2, 1, 2), include.constant=TRUE)

# Summarize Results
models <- data.frame(p = rep(c(0, 0, 0, 1, 1, 1, 2, 2, 2), 2),
                     d = rep(1, 18),
                     q = rep(c(0, 1, 2), 6),
                     include.constant = c(rep(FALSE, 9), rep(TRUE, 9)),
                     loglik = c(fit.00$loglik, fit.01$loglik, fit.02$loglik,
                                fit.10$loglik, fit.11$loglik, fit.12$loglik,
                                fit.20$loglik, fit.21$loglik, fit.22$loglik,
                                fit.00c$loglik, fit.01c$loglik, fit.02c$loglik,
                                fit.10c$loglik, fit.11c$loglik, fit.12c$loglik,
                                fit.20c$loglik, fit.21c$loglik, fit.22c$loglik),
                     aicc = c(fit.00$aicc, fit.01$aicc, fit.02$aicc,
                                fit.10$aicc, fit.11$aicc, fit.12$aicc,
                                fit.20$aicc, fit.21$aicc, fit.22$aicc,
                                fit.00c$aicc, fit.01c$aicc, fit.02c$aicc,
                                fit.10c$aicc, fit.11c$aicc, fit.12c$aicc,
                                fit.20c$aicc, fit.21c$aicc, fit.22c$aicc)
                     )
print(models, digits=6)
```
These results are unreliable.  Use the method described in the next seciton
instead.




AICC for Candidate Models (Correct Version)
-------------------------------------------

```{r}
# Without constant:
fit.00 <- Arima(diff.log.GOOG, c(0, 0, 0), include.constant=FALSE)
fit.01 <- Arima(diff.log.GOOG, c(0, 0, 1), include.constant=FALSE)
fit.02 <- Arima(diff.log.GOOG, c(0, 0, 2), include.constant=FALSE)
fit.10 <- Arima(diff.log.GOOG, c(1, 0, 0), include.constant=FALSE)
fit.11 <- Arima(diff.log.GOOG, c(1, 0, 1), include.constant=FALSE)
fit.12 <- Arima(diff.log.GOOG, c(1, 0, 2), include.constant=FALSE)
fit.20 <- Arima(diff.log.GOOG, c(2, 0, 0), include.constant=FALSE)
fit.21 <- Arima(diff.log.GOOG, c(2, 0, 1), include.constant=FALSE)
fit.22 <- Arima(diff.log.GOOG, c(2, 0, 2), include.constant=FALSE)

# With constant:
fit.00c <- Arima(diff.log.GOOG, c(0, 0, 0), include.constant=TRUE)
fit.01c <- Arima(diff.log.GOOG, c(0, 0, 1), include.constant=TRUE)
fit.02c <- Arima(diff.log.GOOG, c(0, 0, 2), include.constant=TRUE)
fit.10c <- Arima(diff.log.GOOG, c(1, 0, 0), include.constant=TRUE)
fit.11c <- Arima(diff.log.GOOG, c(1, 0, 1), include.constant=TRUE)
fit.12c <- Arima(diff.log.GOOG, c(1, 0, 2), include.constant=TRUE)
fit.20c <- Arima(diff.log.GOOG, c(2, 0, 0), include.constant=TRUE)
fit.21c <- Arima(diff.log.GOOG, c(2, 0, 1), include.constant=TRUE)
fit.22c <- Arima(diff.log.GOOG, c(2, 0, 2), include.constant=TRUE)

# Summarize Results
models <- data.frame(p = rep(c(0, 0, 0, 1, 1, 1, 2, 2, 2), 2),
                     d = rep(1, 18),
                     q = rep(c(0, 1, 2), 6),
                     include.constant = c(rep(FALSE, 9), rep(TRUE, 9)),
                     loglik = c(fit.00$loglik, fit.01$loglik, fit.02$loglik,
                                fit.10$loglik, fit.11$loglik, fit.12$loglik,
                                fit.20$loglik, fit.21$loglik, fit.22$loglik,
                                fit.00c$loglik, fit.01c$loglik, fit.02c$loglik,
                                fit.10c$loglik, fit.11c$loglik, fit.12c$loglik,
                                fit.20c$loglik, fit.21c$loglik, fit.22c$loglik),
                     aicc = c(fit.00$aicc, fit.01$aicc, fit.02$aicc,
                                fit.10$aicc, fit.11$aicc, fit.12$aicc,
                                fit.20$aicc, fit.21$aicc, fit.22$aicc,
                                fit.00c$aicc, fit.01c$aicc, fit.02c$aicc,
                                fit.10c$aicc, fit.11c$aicc, fit.12c$aicc,
                                fit.20c$aicc, fit.21c$aicc, fit.22c$aicc)
                     )
print(models, digits=6)
```

Selected Model
--------------

```{r}
fit.best <- Arima(log.GOOG, c(0, 1, 0), include.constant=TRUE)
print(fit.best)
```

Note: constant (`drift`) is not significant.  However, the hypothesis tests
relies on the model being correct, so the z-statistic and corresponding
p-value are probably not reliable.  It's better to use AICC to determine
whether to include a constant.


Residuals
---------

```{r}
resid <- residuals(fit.best)
plot(time, resid, type="l", col=2)
Acf(resid)
Pacf(resid)
```


Ljung-Box Goodness-of-Fit Test
------------------------------

```{r}
# fitdf = 1 (p = 0, q = 0, model includes constant)
Box.test(resid, lag=12, type = "Ljung-Box", fitdf=1)
Box.test(resid, lag=24, type = "Ljung-Box", fitdf=1)
Box.test(resid, lag=36, type = "Ljung-Box", fitdf=1)
Box.test(resid, lag=48, type = "Ljung-Box", fitdf=1)
```



Forecasts From Selected Model
-----------------------------
```{r}
forecast(fit.best, h=10, level=95)
plot(forecast(fit.best, h=100, level=95), col=2)
```


Forecasts From Model Without Constant
-------------------------------------
```{r}
fit.00 <- Arima(log.GOOG, c(0, 1, 0), include.constant=FALSE)
plot(forecast(fit.00, h=100, level=95), col=2)
```


Don't Use p-values for Model Selection
--------------------------------------

```{r}
fit.11 <- Arima(log.GOOG, c(1, 1, 1), include.constant=FALSE)
print(fit.11)

(zstat.ar1 <- 0.895 / 0.140)
(pval.ar1 <- 2*pnorm(-abs(zstat.ar1)))

(zstat.ma1 <- -0.866 / 0.157)
(pval.ma1 <- 2*pnorm(-abs(zstat.ma1)))
```

