---
title: Forecasting Time Series
subtitle: Seasonal Models
author: Patrick Perry
date: March 24, 2015
---

```{r setup, cache=FALSE, echo=FALSE}
library("RColorBrewer")    # brewer.pal
library("knitr")           # opts_chunk

# terminal output
options(width = 80)

# color palette
palette(brewer.pal(6, "Set1"))

# code chunk options
opts_chunk$set(cache=TRUE, fig.align="center", comment=NA, echo=TRUE,
               tidy=FALSE)
```


```{r}
library("forecast")
```


The dataset is the log unemployment rate, monthly, seasonally adjusted.

```{r}
data <- read.csv("http://ptrckprry.com/course/forecasting/data/unemployment.csv")
date <- as.Date(data$date)
time <- 1:length(date)
unemp <- data$unemployment
log.unemp <- log(unemp)
```

---

Time series plot:

```{r}
plot(date, log.unemp, type="l", col=2)
```

ACF and PACF:

```{r}
par(mfrow=c(1,2))
Acf(log.unemp)
Pacf(log.unemp)
```

---

In HW5, we fit an ARIMA(2,1,2) model without a constant.  The ACF of the
residuals shows a clear seasonal correlation pattern (even though the original
series was supposed to be seasonally adjusted).

```{r}
fit1 <- Arima(log.unemp, c(2,1,2))
resid1 <- residuals(fit1)
Acf(resid1)
```

This is confirmed by the Ljung-Box tests.
```{r}
Box.test(resid1, lag=12, type="Ljung-Box", fitdf=4)
Box.test(resid1, lag=24, type="Ljung-Box", fitdf=4)
Box.test(resid1, lag=36, type="Ljung-Box", fitdf=4)
Box.test(resid1, lag=48, type="Ljung-Box", fitdf=4)
```

---


The original data source claims that it is seasonally adjusted, but it does
not appear that the seasonal trend has been completely removed.  Let's go back
to the original (non-seasonal adjusted) data:

```{r}
data.nsa <- read.csv("http://ptrckprry.com/course/forecasting/data/unemployment-nsa.csv")

unemp.nsa <- data.nsa$unemployment
log.unemp.nsa <- log(unemp.nsa)
```

---

For comparison, here is the original series, along with the
seasonally-adjusted version

```{r}
plot(date, log.unemp.nsa, ylab="Log Unemp.", type="l", col=2)
lines(date, log.unemp, col=1)
```

---

For the rest of this analysis, we will work with the original (not
"seasonally-adjusted" data):

```{r}
data <- read.csv("http://ptrckprry.com/course/forecasting/data/unemployment-nsa.csv")
date <- as.Date(data$date)
time <- 1:length(date)
unemp <- data$unemployment
log.unemp <- log(unemp)
```


---

One simple method of seasonal adjustment is to remove the monthly averages
before analyzing the series.  Here is code to compute these averages and
adjust the series

```{r}
# Extract the month of each observation
month <- factor(months(date), levels=format(ISOdate(2000,1:12,1),"%B"))

# Compute the monthly averages
month.avg <- c(by(log.unemp, month, mean))
print(month.avg)

log.unemp.adj <- log.unemp - month.avg[month]
```

---

For the log unemployment series, the adjustment doesn't work very well:
```{r}
plot(date, log.unemp - mean(log.unemp), type="l", col=2)
lines(date, log.unemp.adj, col=1)

Acf(diff(log.unemp.adj))
```

---

Let's try selecting a seasonal ARIMA model with AICC:

```{r}
d <- 1
D <- 1
for (p in 0:2) {
    for (q in 0:2) {
        for (P in 0:2) {
            for (Q in 0:2) {
                fit <- list(aicc=NA)
                try({
                    fit <- Arima(log.unemp, c(p,d,q),
                                 seasonal=list(order=c(P,D,Q), period=12),
                                 method="ML")
                })
                cat("ARIMA",
                    "(", p, ",", d, ",", q, ")",
                    "(", P, ",", D, ",", Q, ")[12]",
                    " : ", fit$aicc, "\n", sep="")
            }
        }
    }
}
```

---

```{r}
fit <- Arima(log.unemp, c(2,1,2), seasonal=list(order=c(0,1,1), period=12))
resid <- residuals(fit)
par(mfrow=c(1,2))
Acf(resid)
Pacf(resid)
```

```{r}
Box.test(resid, lag=12, type="Ljung-Box", fitdf=5)
Box.test(resid, lag=24, type="Ljung-Box", fitdf=5)
Box.test(resid, lag=36, type="Ljung-Box", fitdf=5)
Box.test(resid, lag=48, type="Ljung-Box", fitdf=5)
```

```{r}
fit0 <- Arima(log.unemp, c(2,1,2))
```

Short-Range (10 Month) Forecasts
--------------------------------

```{r}
n <- max(time)
par(mfrow=c(1,2))
plot(forecast(fit, h=10))
plot(forecast(fit, h=10), xlim=c(n - 18, n + 10))

par(mfrow=c(1,2))
plot(forecast(fit0, h=10))
plot(forecast(fit0, h=10), xlim=c(n - 18, n + 10))
```

Mid-Range (2 Year) Forecasts
----------------------------

```{r}
par(mfrow=c(1,2))
plot(forecast(fit, h=24))
plot(forecast(fit, h=24), xlim=c(n - 100, n + 2*12))

par(mfrow=c(1,2))
plot(forecast(fit0, h=24))
plot(forecast(fit0, h=24), xlim=c(n - 100, n + 2*12))
```

Long-Range (10 Year) Forecasts
------------------------------

```{r}
par(mfrow=c(1,2))
plot(forecast(fit, h=10*12))
plot(forecast(fit, h=10*12), xlim=c(n - 100, n + 10*12))

par(mfrow=c(1,2))
plot(forecast(fit0, h=10*12))
plot(forecast(fit0, h=10*12), xlim=c(n - 100, n + 10*12))
```

Long-Range (50 Year) Forecasts
------------------------------

```{r}
plot(forecast(fit, h=50*12))
plot(forecast(fit0, h=50*12))
```
