Homework 2: Problems 1--3
=========================
*Firstname LastName (Replace this part with your name)*


---


In Problems 1--3, we analyze the Russell 2000 Stock Index, adjusted daily
closing price, recorded from 10 September 1987 to 10 February 2017.

```{r}
data <- read.csv("http://ptrckprry.com/course/forecasting/data/russell.csv")
russell <- data$russell
date <- as.Date(data$date)
time <- 1:length(russell)
```

We will work with the mean-adjusted series:

```{r}
today <- russell - mean(russell)
```


Problem 1
=========

Part A
------

Here is a plot of Today's Russell versus time, as well as Yesterday's Russell
versus time:

```{r}
# The following command creates a lagged version of the 'today' vector.
# It works by dropping the last value of the today vector, and then
# prepending a missing value ('NA'):
#
#     today      x1 x2 x3 x4
#     yesterday  NA x1 x2 x3
#
# By construction, yesterday[t] is equal to today[t-1] for all t > 1.
#
# Note that a negative index value (-length(today), in this instance)
# means "drop the element at this index".
#
yesterday <- c(NA, today[-length(today)])

plot(date, today, xlab="Date", ylab="Mean-Adjusted Russell", type="l", lty=1)
lines(date, yesterday, lty=3) # adds lines for yesterday to the existing plot

legend("topleft", inset=0.05,
       legend=c("Today", "Yesterday"), lty=c(1,3))
```

Next, here is a plot of Today's Russell versus time, as well as
0.5 (Yesterday's Russell):


```{r}
## Replace this comment with code to make the plot
```


Part B
------

**Based on these two plots, which seems to be a better forecast of Today's
Russell: Today's Russell, or 0.5 (Yesterday's Russell)?**



Part C
------

Here are the average squared forecast errors for the two forecasts:

```{r}
# Replace this comment with code to compute average squared forecast
# errors for the two forecasts (you will compute two numbers: one for
# 'yesterday', and one for '0.5 yesterday').
#
# Hint: use the mean function.
```

**Based on this, which one was better?**




Problem 2
=========

Part A
------

Here is a plot of Today's Russell versus Yesterday's Russell:

```{r}
# Add code to make the plot.  Put "Today's Russell" on the y-axis.
```

**Describe any patterns you see.**



Part B
------

Here is a linear regression fit of Today's Russell on Yesterday's Russell:

```{r}
# Replace the ???? with code to fit the linear regression model, then
# uncomment the following two lines.
#
# model <- ????
# summary(model)
```

**What is the prediction of Today's Russell implied by the regression
coefficients?**


**Is this consistent with your answers to Problem 1, parts B and C?**



Part C
------

The 95% confidence interval for the slope is
```{r}
# Add code to compute the confidence interval for the coefficient of
# "yesterday".
#
# Hint: use the confint function.
```

**Is the slope in your fitted regression significantly different from 1?**


The 95% confidence interval for the intercept is
```{r}
# Add code to compute the confidence interval for the intercept.
# Hint: use the confint function
```

**Briefly comment on the intercept as well.**


Part D
------
**Based on everything you have done so far, do you see any strong evidence
that the Russell is *not* a random walk**?



Part E
------
Here is the correlation coefficient between Today's Russell and Yesterday's
Russell:

```{r}
# There are two ways to do this in R:  
#
# 1. You can get the R^2 with the command `summary(model)$r.squared`,
#    then take the square root, and then multiply by the sign of the
#    coefficient, `sign(coef(model)[["yesterday"]])`.
#
# 2. You can use the `cor` function.
# 
# Choose either method and compute the value here.
```


**Based on this, how strong is the linear association between Today's Russell
and Yesterday's Russell?**



Problem 3
=========

Part A
------

We define the Russell returns as
```{r}
# Replace the ???? with code to compute the Russell returns, then
# uncomment the following two lines:
# ret.today <- ????
# ret.yesterday <- c(NA, ret.today[-length(ret.today)])
```


Here is a plot of Today's return versus time:
```{r}
# Add the plot here.
```


Here are the sample average and standard deviation of the returns:
```{r}
# Hint: use the mean and sd functions.
```


Here is an ordinary t-test of whether the mean returns are zero:
```{r}
# Replace this comment with code to perform a t-test on the data
# stored in `ret.today`.
#
# Hint: use the t.test function.
```

**Are the mean returns significantly different from zero?  Interpret your
findings.**



Part B
------

Here is a histogram of the Russell returns:
```{r}
# Add code to make the histogram.
#
# Hint: use the hist command.
```

Here is a boxplot:
```{r}
# Add code to make the boxplot.
#
# Hint: use the boxplot command.
```

Finally, here is a normal quantile-quantile plot:
```{r}
# Add code to make the normal quantile-quantile plot.
# 
# Hint: use the qqnorm command.
```

**Do you think that the Russell returns are normally distributed?  Explain.**



Part C
------

Here is a plot of today's return versus yesterday's return
```{r}
# Add code to make the plot.
```

Here is a zoomed in plot:
```{r}
# Due to the extreme values, there is not much information in the
# plot above.  You can zoom in on the plot by adding xlim and
# ylim arguments, as in plot(x, y, xlim=c(1, 3), ylim=c(-2,2)); this would
# sets the x limits to [1,3] and sets the y limits to [-2,2].

# Replace this comment by taking the previous plot command and then
# adding appropriate values for xlim and ylim.
```


**Does this plot appear very different from the one in Problem 2(A)?  Which
seems to be easier to predict: Today's Russell, or Today's return?**



Part D
------

Here is a linear regression of today's returns on yesterday's returns:
```{r}
# Add code to fit the model and print a summary of the fit.
```

**What is the prediction of today's return implied by the regression
coefficients?**


**Are the coefficients statistically significantly different from zero?**

